Building Markov State Models for Periodically Driven Non-Equilibrium Systems

### 2015-02-09 writedown/period.noneq.msm/submit.jctc.reply/paper.tex
backup

Biomolecular systems under non-equilibrium conditions caused by external fields, especially
systems under periodic forcing, have attracted increasing
interest recently.  For example, the potential effects of electromagnetic
radiation on  human body tissues (e.g.,~DNA, protein, and membrane)  have been extensively
investigated in a vast number of studies, with the following list  representing an incomplete selection~\cite{bohr2000microwave, bohr2000microwave-1, dePomerai2000cell,
  dePomerai2003microwave, mancinelli2004non, inskip2001cellular, bekard2013electric, budi2005electric,
  budi2007effect, budi2008comparative, astrakas2012structural,
  damm2012can, english2009nonequilibrium, solomentsev2012effects}.  Molecular dynamics (MD) simulations have proven particularly useful for understanding
the response of biomolecular conformations to external fields, because of their ability to  resolve molecular
details that sometimes cannot be resolved in experiments. Only
recently, a non-equilibrium MD simulation (D-NEMD) method with local thermostating has been proposed~\cite{wang2014exploring} that allows for studying 
non-equilibrium processes in a statistically reliable and thermodynamically consistent way. 
Despite the significance of the non-equilibrium phenomena, 
analysis of the non-equilibrium MD simulations mainly follows standard approaches, and reliable tools for quantitative description of
the essential conformational dynamics of the molecular system under external forcing remain unavailable.

@@

Despite their many advantages, MD simulations have severe limitations. For example, one has to assume that the underlying force fields  describe the internal and external molecular interactions appropriately, and the maximal possible simulation length often is shorter than the timescale of interest. 
This article is mainly concerned with circumventing the latter obstacle by introducing non-equilibrium Markov State Models. Markov State Models
(MSM) have been  developed thoroughly over the
past decade in theory~\cite{A19-31,prinz2011markov} and applications (see the recent book~\cite{A19-1} for an overview), and software implementations~\cite{A19-49, MSMBuilder}, but for systems under equilibrium conditions only.
The main concept underlying equilibrium MSMs is to approximate the original high-dimensional MD system 
using reduced Markovian dynamics over a finite number of (macro-)states. These (macro-)states
have to be identified with the dominant metastable sets, in the sense that
typical MD trajectories remain close to a metastable set substantially longer than
they need for a transition to another such state~\cite{A19-31,schuette2011markov}. In this case, the metastable sets are the main conformations of the molecular system under consideration, which, often enough, are given by the main wells in the energy landscape. 
It has been shown that for molecular systems exhibiting such metastable sets, the Markovian dynamics given by an MSM allows very close approximation of the longest relaxation processes of the underlying molecular system, at least under equilibrium conditions~\cite{sarich2010approximation,Eigenvalues}. 
Moreover, it has been demonstrated that, in such cases, MSM building requires short MD trajectories only, which are much shorter than the timescales of interest~\cite{PNAS09,kohlhoff2014cloud}.  
Thus, MSM building often allows the study of dynamic behavior on long timescales without requiring MD trajectories of comparable length.
Moreover, MSMs are utilized for understanding very long MD simulations. Extracting the essential structures and dynamic properties from long MD runs is
becoming increasingly difficult as the system size and trajectory length grow; MSMs have been used to construct kinetic fingerprints from MD simulations~\cite{A19-39} which facilitates understanding of essential dynamics and comparison with experimental data~\cite{PrinzKellerNoe_PCCP11_Perspective}.

@@

% \textbf{Importance: first application of MSM in a non-equilibrium system.}
To the best of our knowledge, this work is the first attempt of using MSMs for analyzing non-equilibrium systems under periodic external forcing.
More precisely, we will demonstrate how to use MSMs to investigate the conformational dynamics of a peptide (alanine dipeptide) under an oscillating electric field (EF).
To this end, we will show how to generalize Markov state modeling to periodic non-equilibrium conditions, where one cannot assume reversibility of the dynamics as is most often case in the literature on MSM building.

The outline of the article is as follows: 
In Section~\ref{sec:disc}, we discuss the temporal and spatial
discretizations needed to construct an MSM. We consider spatial discretization
of the dihedral angle space in the traditional sense of full partition MSMs~\cite{pande2010everything,A19-29}.
In the temporal direction, the non-equilibrium process
is discretized utilizing Floquet's theorem. This results in a time-homogeneous, 
but not necessarily reversible, that is, irreversible Markov process. Since  full spatial partition is not rational
for high-dimensional systems, we next show how to construct a \emph{few-states} MSM, based on milestoning~\cite{schuette2011markov,A19-29} of the discretized irreversible Markov process in Section~\ref{sec:build-msm}.
The validity of the discretizations and the resulting MSM is checked in Section~\ref{sec:alanine}
by comparing the kinetic fingerprints given by the MSM to brute force
non-equilibrium MD simulations of alanine dipeptide under oscillating EF.
The findings are summarized in Section~\ref{sec:conclusion}, which includes a chart presenting
the workflow of MSM building for non-equilibrium systems, and a list of open questions.

@@

\section{Non-equilibrium molecular dynamics and its discretization}
\label{sec:disc}
We consider diffusive molecular dynamics in an energy landscape $V$, driven by the time-dependent external driving force $E(t)D(x_t)$, with the
$T$-periodic external field $E(t)$:
\begin{align}
  \label{eq:disc-1}
  d\vect x_t = \Big(-\nabla V(\vect x_t) + E(t) D(\vect x_t)\Big)dt + \sqrt{2\beta^{-1}} d\vect w_t, 
\end{align}
where $x_t\in\Omega$ denotes the state of the molecular system at time $t$ in state space $\Omega$, $\vect w_t$ denotes standard $n$-dimensional Brownian motion, and $\beta$ the inverse temperature,
i.e.,~$\beta = 1/(k_B\mathcal T)$.
Thermostatted Hamiltonian or Langevin dynamics can be treated in the same way as explained herein, so for the sake of simplicity, we focus on the discussion of diffusive dynamics.
The propagation of probability
densities $\rho=\rho(\vect x,t)$ based on this kind of dynamics in the sense
of $\rho(\vect x,t)dx=\prob[\vect x_t\in [\vect x,\vect x+d\vect x)]$, is governed by
the Fokker-Planck equation:
\begin{align}
  \label{eq:disc-fp}
  \frac{\partial \rho}{\partial t} = \ml^\dagger(t) \rho,
\end{align}
where $\ml^\dagger(t)$ is the adjoint of the generator
\begin{align}
  \label{eq:disc-3}
  \ml(t)=\beta^{-1}\Delta_{\vect x}+\Big(-\nabla_x V(\vect x) + E(t)D(\vect x)\Big)\cdot\nabla_{\vect x},
\end{align}
where $\Delta_{\vect x}$ denotes the Laplacian operator and $\nabla_{\vect x}$
the nabla-operator with respect to to $\vect x$. 
The periodicity of the external driving force induces the periodicity of the generator,
i.e.,~$\ml(t) = \ml(t+T)$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Spatial discretization: Master equation}
We will now introduce an appropriate spatial discretization of this kind of non-equilibrium MD --  this is done for reasons of simplicity only; we could completely avoid it for the price of more technical arguments.
For achieving this discretization, we 
introduce a partition of state space $\Omega$ into a finite number of disjoint
sets $\{ \Omega_1, \cdots, \Omega_N\}$ satisfying $\Omega = \cup_i \Omega_i$,
$\Omega_j\cap \Omega_j = \emptyset,\ \forall i\neq j$.
Utilizing the procedure described in Ref.~\cite{latorre2011structure} the original Fokker-Planck equation~(\ref{eq:disc-fp})
is discretized, resulting in a time-inhomogeneous Markov jump process in state
space $S = \{1, \cdots, N\}$, with a time-dependent rate
matrix $\vect L(t) \in \mathbb R^{N\times N}$ satisfying
\begin{align}\label{eq:disc-4}
\sum\limits_{j=1}^N L_{ij}(t) & =  0\\ \label{eq:disc-5}
L_{ij}(t) & \ge  0, \quad i\not= j\\
L_{ij}(t) & =  L_{ij}(t+T)
\end{align}
for all real time $t\geq 0$.
Moreover, the rate matrix $L$ has the form $\vect L(t)=\vect L_0+E(t)\vect L_1$,
where $E(t)$ is periodic, with period $T>0$.
In analogy to~\eqref{eq:disc-fp}, the Markov jump process generated by
$\vect L(t)$ transports probability distributions according to the associated Master equation
\begin{align}
  \label{eq:disc-master}
  \frac{d\vect p(t)}{dt} = \vect L^{\top}(t)\cdot \vect p(t),
\end{align}
where $\vect L^{\top}(t)$ denotes the matrix transpose of $\vect L(t)$. $\vect p(t)$ is an $N$-vector denoting the probability distribution on $S$ at time $t$; $p(i,t)$, for example, is the probability of being in state $i$ (which corresponds to set $\Omega_i$) at time $t$.
As usual,  properties (\ref{eq:disc-4}) and (\ref{eq:disc-5}) of
$\vect L(t)$ guarantee that the total probability mass is conserved,
i.e., if $p(i,0)\ge 0$ component-wise, then $p(i,t)\ge 0$ and $\sum_i
p(i,t) = \sum_ip(i,0)$.
% The solution of the master equation need no be
% periodic.
The temporal evolution of the probability distribution $\vect p(t)$ can be formally written as
\begin{align}  \label{eq:disc-8}
\vect p(t)=\myphi(t)\vect p(0)
\end{align}
by using the
associated propagator matrix $\myphi(t)\in\mathbb R^{N\times N}$ that solves
\begin{align}
  \label{eq:disc-master-phi}
  \frac{d}{dt}\myphi(t) = \vect L^{\top}(t)\myphi(t), \quad \myphi(0) = \id .
\end{align}
Since the last equation can be considered column-wise, the propagator matrix inherits column-wise conservation properties:
$\Phi_{ij}(t) \ge  0$
and $\sum\limits_{i=1}^N \Phi_{ij}(t)  =  1$,
that is, $\myphi^{\top}(t)$ is a
stochastic matrix satisfying $\myphi^{\top}(t)\vect e=\vect e$,
with $\vect e=(1,\ldots,1)^{\top}\in \mathbb R^N$.
Regarding these considerations, we find
\begin{align}
\label{eq:disc-10}  
\myphi_{ij}(t)=\prob\left(\vect X_t=i\mid \vect X_0=j \right),
\end{align}
where $\vect X_t$ denotes the Markov process generated by $\vect L(t)$.

@@

% \subsection{Building the Markov State Model from the Floquet transition matrix}
If the discretization cells $\Omega_i$, $i=1,\ldots,N$, form a fine
partition of the molecular state space, the Markov chain defined via
the transition matrix $P$
 is discrete in time, but in space it remains a fine-scale description of the transport properties of the
dynamics with a very large number ($N$) of states.  Now we want to coarse our description much further by
constructing a Markov State Model (MSM) for $\vect P$ with $K\ll N$
\emph{macrostates}, which would be the metastable states of the system. The resulting $K\times K$ MSM transition matrix $\hat{\vect P}$
then defines the coarse-grained long-term kinetics that shall
approximate the original long term kinetics well. 
The idea behind MSM building is that, given that the molecular system under consideration exhibits metastable conformations,  it is usually possible to construct a
relatively small number of discrete sets --the metastable sets that form the so-called macrostates-- that
correctly describe the slow  dynamics, and 
in each set the fast dynamics relaxes on some timescales that are significantly shorter than the metastable timescales.
Then, if the MSM dynamics reproduces the slow timescales and the corresponding transitions
of the original dynamics~\eqref{eq:disc-1},
the former is considered to be a good approximation of the latter.

@@

However, since we cannot assume $P$ to satisfy such a detailed balance, we
instead follow the approach for non-standard MSM construction recently proposed in Ref.~\cite{sarich2014utilizing}, which allows  identification of the metastable core sets for the
non-reversible transition matrix $P$. Assume that this approach leads
to the $K$ core sets $C_1,\ldots, C_K\subset S$ that are appropriate metastable sets. 
Following~\cite{A19-31,schuette2011markov}, the process $(\tilde X_m)$ associated with $P$ is coarse-grained into the so-called milestone process $(\hat{X}_m)$ in the following way: 
\begin{itemize}
\item $\hat{X}_m$ only has $K$ states associated with the sets $C_j$, $j=1,\ldots,K$. 
\item The sequence of random variables $(\hat{X}_m)$ is defined via the sequence $(\tilde X_m)$, i.e., trajectories of $(\tilde X_m)$ induce trajectories of $(\hat{X}_m)$. We set $\hat{X}_m=j$ if the last core set that the process $(\tilde X_m)$ entered prior to or at time $m$ is the core set $C_j$.
\end{itemize}
Now, consider an arbitrary infinitely long trajectory of $(\tilde X_m)$. Because of ergodicity, we know that the states in this trajectory will be distributed according to the quasi-stationary distribution $\mu$. Based on such an infinitely long trajectory, we can consider the probability $q_j^\bwd(i)$ that conditioned on $\tilde X_m=i$ the last core set hit has been $C_j$.
This function is called the backward committor of $(\tilde X_m)$ associated with the set $C_j$ and is associated with the milestone process via
\begin{align}
\label{eq:msm-bwd-committor}  
q^\bwd_j(i)=\prob_\mu(\hat{X}_m=j \mid X_m=i),
\end{align}
where  the index $\mu$ refers to the fact that $\hat{X}_m$ is distributed due to $\mu$.
%\recheck{\underline{QUESTION}: Here it is a \emph{conditional} probability,
%  due to the time-homogeneity,
%  it does not matter whether it is defined from quasi-stationary probability
%  or not.}
From the last equation, we get that the stationary distribution of the milestone process is given by
\begin{align}  
\hat{\mu}_j=\sum_{i\in S} q^\bwd_j(i)\mu(i),
\end{align}
that is, the probability to find $\hat{X}_m=j$ in (infinitely) long trajectories of the milestoning process is $\hat{\mu}_j$.

@@

We define the one-step transition matrix $\hat{\vect P}$ for the milestone process by
\begin{equation}\label{hatP_prob}
\hat{\vect P}_{jk}=\prob_\mu\Big(\hat{X}_{m+1}=k\,\vert\, \hat{X}_m=j\Big).
\end{equation}
Then, following
Ref.~\cite{djurdjevac2010markov}, Thm.~3.1, 
$\hat{\vect P}$ can be computed by matrix multiplication using the committors:
  \begin{align}
    \label{eq:msm-tmatrix-00}
    \begin{split}      
    \hat{\vect P}_{jk}
    = &
    \frac{1}{\hat{\mu}_j}
    \langle (\vect P^b - \id) q^\bwd_j,q^\fwd_k \rangle_\mu,\qquad j\not= k, \\    %\label{eq:msm-tmatrix-01}
    \hat{\vect P}_{jj}
    =&
    1-\sum_{k\not=j} \hat{\vect P}_{jk}
    \end{split}
  \end{align}
where the inner product is defined by
$\langle u,v \rangle_\mu=\sum_{i\in S} u(i) v(i) \mu(i)$. 
In general, the milestone process need not be a Markov process. The results in \cite{A19-31,sarich2014utilizing} show, however, that it is an approximate Markov process as long as the core sets are proper metastable sets, i.e., if the typical timescale on which $(X_m)$ leaves $C$ is much smaller than the typical expected hitting times between the core sets.
Thus, by taking $\hat{P}$ as our MSM transition matrix, we introduce an additional modeling error that is the smaller the more metastable the core sets are.
With this MSM transition matrix, we can define the MSM kinetics:
If we start from some initial probability $\hat{p}_j(0)$ of being in state $j$ at time $t=0$ then its evolution $\hat{p}_j(t)$ in time is discrete in multiples of period $T$ and given by
\begin{equation}\label{eq:num-29}
\hat{p}_j(T)=\sum_k \hat{p}_k(0)\hat{\vect P}_{kj}.
\end{equation}
In our approach $\hat p_j(mT)$ is a good approximation of $\prob(\hat{X}_m=j)$ (for appropriately chosen core sets).

@@

For this periodically driven molecular system, we will first show how to choose an appropriately fine spatial discretization. After validating this discretization, we will consider
the time-discretized dynamics generated by the Floquet transition matrix $\vect P$
in comparison to the original NEMD simulation. Finally, we will coarse grain this description further by construction of a $3$-state Markov State Model that is able to describe the long-term kinetics of the system correctly.

>>>>

Biomolecular systems under non-equilibrium conditions caused by external fields, especially
systems under periodic forcing, have attracted increasing
interest recently.  For example, the potential effects of electromagnetic
radiation on  human body tissues (e.g.,~DNA, protein, and membrane)  have been extensively
investigated in a vast number of studies, with the following list  representing an incomplete selection Ref.~\cite{bohr2000microwave, bohr2000microwave-1, dePomerai2000cell,
  dePomerai2003microwave, mancinelli2004non, inskip2001cellular, bekard2013electric, budi2005electric,
  budi2007effect, budi2008comparative, astrakas2012structural,
  damm2012can, english2009nonequilibrium, solomentsev2012effects}.  Molecular dynamics (MD) simulations have proven particularly useful for understanding
the response of biomolecular conformations to external fields, because of their ability to  resolve molecular
details that sometimes cannot be resolved in experiments. Only
recently, a non-equilibrium MD simulation (NEMD) method with local thermostating has been proposed~\cite{wang2014exploring} that allows for studying 
non-equilibrium processes in a statistically reliable and thermodynamically consistent way. 
Despite the significance of the non-equilibrium phenomena, 
analysis of the non-equilibrium MD simulations mainly follows standard approaches, and reliable tools for quantitative description of
the essential conformational dynamics of the molecular system under external forcing remain unavailable.

@@

Despite their many advantages, MD simulations have severe limitations. For example, one has to assume that the underlying force fields  describe the internal and external molecular interactions appropriately, and the maximal possible simulation length often is shorter than the timescale of interest. 
This article is mainly concerned with circumventing the latter obstacle by introducing non-equilibrium Markov State Models. Markov State Models
(MSM) have been  developed thoroughly over the
past decade in theory~\cite{A19-31,prinz2011markov} and applications (see the recent book~\cite{A19-1} for an overview), and software implementations~\cite{A19-49, MSMBuilder}, but for systems under equilibrium conditions only.
The main concept underlying equilibrium MSMs is to approximate the original high-dimensional MD system 
using reduced Markovian dynamics over a finite number of (macro-)states. These (macro-)states
have to be identified with the dominant metastable sets, in the sense that
typical MD trajectories remain close to a metastable set substantially longer than
they need for a transition to another such set~\cite{A19-31,schuette2011markov}. In this case, the metastable sets are the main conformations of the molecular system under consideration, which, often enough, are given by the main wells in the energy landscape. 
It has been shown that for molecular systems exhibiting such metastable sets, the Markovian dynamics given by an MSM allows very close approximation of the longest relaxation processes of the underlying molecular system, at least under equilibrium conditions~\cite{sarich2010approximation,Eigenvalues}. 
Moreover, it has been demonstrated that, in such cases, MSM building requires short MD trajectories only, which are much shorter than the timescales of interest~\cite{PNAS09,kohlhoff2014cloud}.  
Thus, MSM building often allows the study of dynamic behavior on long timescales without requiring MD trajectories of comparable length.
Moreover, MSMs are utilized for understanding very long MD simulations. Extracting the essential structures and dynamic properties from long MD runs is
becoming increasingly difficult as the system size and trajectory length grow; MSMs have been used to construct kinetic fingerprints from MD simulations~\cite{A19-39} which facilitates understanding of essential dynamics and comparison with experimental data~\cite{PrinzKellerNoe_PCCP11_Perspective}.

@@

% \textbf{Importance: first application of MSM in a non-equilibrium system.}
To the best of our knowledge, this work is the first attempt of using MSMs for analyzing non-equilibrium systems under periodic external forcing.
More precisely, we will demonstrate how to use MSMs to investigate the conformational dynamics of a peptide (alanine dipeptide) under an oscillating electric field (EF).
To this end, we will show how to generalize Markov state modeling to periodic non-equilibrium conditions, where one cannot assume reversibility of the dynamics as is most often the case in the literature on MSM building.

The outline of the article is as follows: 
In Section~\ref{sec:disc}, we discuss the temporal and spatial
discretizations needed to construct an MSM. We consider spatial discretization
of the dihedral angle space in the traditional sense of full partition MSMs~\cite{pande2010everything,A19-29}.
In the temporal direction, the non-equilibrium process
is discretized utilizing Floquet's theorem. This results in a time-homogeneous, 
but not necessarily reversible, that is, irreversible Markov process. Since  full spatial partition is not rational
for high-dimensional systems, we next show how to construct a \emph{few-states} MSM, based on milestoning~\cite{schuette2011markov,A19-29} of the discretized irreversible Markov process in Section~\ref{sec:build-msm}.
The validity of the discretizations and the resulting MSM is checked in Section~\ref{sec:alanine}
by comparing the kinetic fingerprints given by the MSM to brute force
NEMD simulations of alanine dipeptide under oscillating EF.
The findings are summarized in Section~\ref{sec:conclusion}, which includes a chart presenting
the workflow of MSM building for non-equilibrium systems, and a list of open questions.

@@

\section{Non-equilibrium molecular dynamics and its discretization}
\label{sec:disc}
We consider diffusive molecular dynamics in an energy landscape $V$, driven by the time-dependent external driving force $E(t)D(x_t)$, with the
$T$-periodic external field $E(t)$:
\begin{align}
  \label{eq:disc-1}
  d\vect x_t = \Big(-\nabla V(\vect x_t) + E(t) D(\vect x_t)\Big)dt + \sqrt{2\beta^{-1}} d\vect w_t, 
\end{align}
where $x_t\in\Omega$ denotes the state of the molecular system at time $t$ in state space $\Omega \subset \mathbb R^{3N_{\textrm{atom}}}$ ($N_{\textrm{atom}}$ being the number of atoms), $\vect w_t$ denotes standard $3N_{\textrm{atom}}$-dimensional Brownian motion, and $\beta$ the inverse temperature,
i.e.,~$\beta = 1/(k_B\mathcal T)$.
Thermostatted Hamiltonian or Langevin dynamics can be treated in the same way as explained herein, so for the sake of simplicity, we focus on the discussion of diffusive dynamics.
The propagation of probability
densities $\rho=\rho(\vect x,t)$ based on this kind of dynamics in the sense
of $\rho(\vect x,t)dx=\prob[\vect x_t\in [\vect x,\vect x+d\vect x)]$, is governed by
the Fokker-Planck equation:
\begin{align}
  \label{eq:disc-fp}
  \frac{\partial \rho}{\partial t} = \ml^\dagger(t) \rho,
\end{align}
where $\ml^\dagger(t)$ is the adjoint of the generator
\begin{align}
  \label{eq:disc-3}
  \ml(t)=\beta^{-1}\Delta_{\vect x}+\Big(-\nabla_x V(\vect x) + E(t)D(\vect x)\Big)\cdot\nabla_{\vect x},
\end{align}
where $\Delta_{\vect x}$ denotes the Laplacian operator and $\nabla_{\vect x}$
the nabla-operator with respect to $\vect x$. 
The periodicity of the external driving force induces the periodicity of the generator,
i.e.,~$\ml(t) = \ml(t+T)$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Spatial discretization: Master equation}
We will now introduce an appropriate spatial discretization of this kind of NEMD --  this is done for reasons of simplicity only; we could completely avoid it for the price of more technical arguments.
For achieving this discretization, we 
introduce a partition of state space $\Omega$ into a finite number of disjoint
sets $\{ \Omega_1, \cdots, \Omega_N\}$ satisfying $\Omega = \cup_i \Omega_i$,
$\Omega_j\cap \Omega_j = \emptyset,\ \forall i\neq j$.
Utilizing the procedure described in Ref.~\cite{latorre2011structure} the original Fokker-Planck equation~(\ref{eq:disc-fp})
is discretized, resulting in a time-inhomogeneous Markov jump process in state
space $S = \{1, \cdots, N\}$, with a time-dependent rate
matrix $\vect L(t) \in \mathbb R^{N\times N}$ satisfying
\begin{align}\label{eq:disc-4}
\sum\limits_{j=1}^N L_{ij}(t) & =  0\\ \label{eq:disc-5}
L_{ij}(t) & \ge  0, \quad i\not= j\\
L_{ij}(t) & =  L_{ij}(t+T)
\end{align}
for all real time $t\geq 0$.
Moreover, the rate matrix $L$ has the form $\vect L(t)=\vect L_0+E(t)\vect L_1$,
where $E(t)$ is periodic, with period $T>0$.
In analogy to~\eqref{eq:disc-fp}, the Markov jump process generated by
$\vect L(t)$ transports probability distributions according to the associated Master equation
\begin{align}
  \label{eq:disc-master}
  \frac{d\vect p(t)}{dt} = \vect L^{\top}(t)\cdot \vect p(t),
\end{align}
where $\vect L^{\top}(t)$ denotes the matrix transpose of $\vect L(t)$. $\vect p(t)$ is an $N$-vector denoting the probability distribution on $S$ at time $t$; $p(i,t)$, for example, is the probability of being in state $i$ (which corresponds to set $\Omega_i$) at time $t$.
As usual,  properties (\ref{eq:disc-4}) and (\ref{eq:disc-5}) of
$\vect L(t)$ guarantee that the total probability mass is conserved,
i.e., if $p(i,0)\ge 0$ component-wise, then $p(i,t)\ge 0$ and $\sum_i
p(i,t) = \sum_ip(i,0)$.
% The solution of the master equation need no be
% periodic.
The temporal evolution of the probability distribution $\vect p(t)$ can be formally written as
\begin{align}  \label{eq:disc-8}
\vect p(t)=\myphi(t)\vect p(0)
\end{align}
by using the
associated propagator matrix $\myphi(t)\in\mathbb R^{N\times N}$ that solves
\begin{align}
  \label{eq:disc-master-phi}
  \frac{d}{dt}\myphi(t) = \vect L^{\top}(t)\myphi(t), \quad \myphi(0) = \id .
\end{align}
Since the last equation can be considered column-wise, the propagator matrix inherits column-wise conservation properties:
$\Phi_{ij}(t) \ge  0$
and $\sum\limits_{i=1}^N \Phi_{ij}(t)  =  1$,
that is, $\myphi^{\top}(t)$ is a
stochastic matrix satisfying $\myphi^{\top}(t)\vect e=\vect e$,
with $\vect e=(1,\ldots,1)^{\top}\in \mathbb R^N$.
Regarding these considerations, we find
\begin{align}
\label{eq:disc-10}  
\myphi_{ij}(t)=\prob\left(\vect X_t=i\mid \vect X_0=j \right),
\end{align}
where $\vect X_t$ denotes the Markov process generated by $\vect L(t)$.

@@

% \subsection{Building the Markov State Model from the Floquet transition matrix}
If the discretization cells $\Omega_i$, $i=1,\ldots,N$, form a fine
partition of the molecular state space, the Markov chain defined via
the transition matrix $P$
 is discrete in time, but in space it remains a fine-scale description of the transport properties of the
dynamics with a very large number ($N$) of states.  Now we want to coarse our description much further by
constructing a Markov State Model (MSM) for $\vect P$ with $K\ll N$
\emph{macrostates}, which would be the metastable states of the system. The resulting $K\times K$ MSM transition matrix $\hat{\vect P}$
then defines the coarse-grained long-term kinetics that shall
approximate the original long-term kinetics well. 
The idea behind MSM building is that, given that the molecular system under consideration exhibits metastable conformations,  it is usually possible to construct a
relatively small number of discrete sets --the metastable sets that form the so-called macrostates-- that
correctly describe the slow  dynamics, and 
in each set the fast dynamics relaxes on some timescales that are significantly shorter than the metastable timescales.
Then, if the MSM dynamics reproduces the slow timescales and the corresponding transitions
of the original dynamics~\eqref{eq:disc-1},
the former is considered to be a good approximation of the latter.

@@

However, since we cannot assume $P$ to satisfy such a detailed balance, we
instead follow the approach for non-standard MSM construction recently proposed in Ref.~\cite{sarich2014utilizing}, which allows  identification of the metastable core sets for the
non-reversible transition matrix $P$. Assume that this approach leads
to the $K$ core sets $C_1,\ldots, C_K\subset S$ that are appropriate metastable sets. 
Following~\cite{A19-31,schuette2011markov}, the process $(\tilde X_m)$ associated with $P$ is coarse-grained into the so-called milestone process $(\hat{X}_m)$ in the following way: 
\begin{itemize}
\item $\hat{X}_m$ only has $K$ states associated with the sets $C_j$, $j=1,\ldots,K$. 
\item The sequence of random variables $(\hat{X}_m)$ is defined via the sequence $(\tilde X_m)$, i.e., trajectories of $(\tilde X_m)$ induce trajectories of $(\hat{X}_m)$. We set $\hat{X}_m=j$ if the last core set that the process $(\tilde X_m)$ entered prior to or at time $m$ is the core set $C_j$.
\end{itemize}
Now, consider an arbitrary infinitely long trajectory of $(\tilde X_m)$. Because of ergodicity, we know that the states in this trajectory will be distributed according to the quasi-stationary distribution $\mu$. Based on such an infinitely long trajectory, we can consider the probability $q_j^\bwd(i)$ that conditioned on $\tilde X_m=i$ the last core set hit has been $C_j$.
This function is called the backward committor of $(\tilde X_m)$ associated with the set $C_j$ and is associated with the milestone process via
\begin{align}
\label{eq:msm-bwd-committor}  
q^\bwd_j(i)=\prob_\mu(\hat{X}_m=j \mid \tilde X_m=i),
\end{align}
where  the index $\mu$ refers to the fact that $\hat{X}_m$ is distributed due to $\mu$.
%\recheck{\underline{QUESTION}: Here it is a \emph{conditional} probability,
%  due to the time-homogeneity,
%  it does not matter whether it is defined from quasi-stationary probability
%  or not.}
From the last equation, we get that the stationary distribution of the milestone process is given by
\begin{align}  
\hat{\mu}_j=\sum_{i\in S} q^\bwd_j(i)\mu(i),
\end{align}
that is, the probability to find $\hat{X}_m=j$ in (infinitely) long trajectories of the milestoning process is $\hat{\mu}_j$.

@@

We define the one-step transition matrix $\hat{\vect P}$ for the milestone process by
\begin{equation}\label{hatP_prob}
\hat{\vect P}_{jk}=\prob_\mu\Big(\hat{X}_{m+1}=k\,\vert\, \hat{X}_m=j\Big).
\end{equation}
Then, following
Ref.~\cite{djurdjevac2010markov}, Thm.~3.1, 
$\hat{\vect P}$ can be computed by matrix multiplication using the committors:
  \begin{align}
    \label{eq:msm-tmatrix-00}
    \begin{split}      
    \hat{\vect P}_{jk}
    = &
    \frac{1}{\hat{\mu}_j}
    \langle (\vect P^b - \id) q^\bwd_j,q^\fwd_k \rangle_\mu,\qquad j\not= k, \\    %\label{eq:msm-tmatrix-01}
    \hat{\vect P}_{jj}
    =&
    1-\sum_{k\not=j} \hat{\vect P}_{jk}
    \end{split}
  \end{align}
where the inner product is defined by
$\langle u,v \rangle_\mu=\sum_{i\in S} u(i) v(i) \mu(i)$. 
In general, the milestone process need not be a Markov process. The results in Ref.~\cite{A19-31,sarich2014utilizing} show, however, that it is an approximate Markov process as long as the core sets are proper metastable sets, i.e., if the typical timescale on which $(X_m)$ leaves $C$ is much smaller than the typical expected hitting times between the core sets.
Thus, by taking $\hat{P}$ as our MSM transition matrix, we introduce an additional modeling error that is the smaller the more metastable the core sets are.
With this MSM transition matrix, we can define the MSM kinetics:
If we start from some initial probability $\hat{p}_j(0)$ of being in state $j$ at time $t=0$, then its evolution $\hat{p}_j(t)$ in time is discrete in multiples of period $T$ and given by
\begin{equation}\label{eq:num-29}
\hat{p}_j(T)=\sum_k \hat{p}_k(0)\hat{\vect P}_{kj}.
\end{equation}
In our approach $\hat p_j(mT)$ is a good approximation of $\prob(\hat{X}_m=j)$ (for appropriately chosen core sets).

@@

For this periodically driven molecular system, we will first show how to choose an appropriately fine spatial discretization. After validating this discretization, we will consider
the time-discretized dynamics generated by the Floquet transition matrix $\vect P$
in comparison to the original NEMD simulation. Finally, we will coarse-grain this description further by construction of a $3$-state Markov State Model that is able to describe the long-term kinetics of the system correctly.
